### 2010-12-13 docs/tex/mimicry.tex
Results chapter completed...
Chapter labels added...
Revision updates...

%Chapter describing the mimetic process
\chapter{The Inspiration: Mimicry}

\begin{quote}
\textsl{``It is not the strongest of the species that survives, nor the most intelligent that survives. It is the one that is the most adaptable to change." - Charles Darwin}
\end{quote}

@@

\begin{itemize}
	\item \textsl{``that the imitative species occur in the same area and occupy the same station as the imitated".} This increases the probability that a predator will affect both parties concerned. Nevertheless, the model could live in Africa and the mimic in Europe, or vice versa, if the deceptive predator were a migratory bird.
	\item \textsl{``that the imitators are always the more defenseless".} On the contrary, for the case of Mertensian mimicry among coral snakes the more offensive party may occasionally be the mimic.
	\item \textsl{``that the imitators are always less numerous in individuals".} What is meant here is that the deceived animal should encounter the mimic less often than the model. This is true when the deceived animal must learn to recognize model and mimic and when positive and negative experiences carry equal weight. But it is known from various learning experiments that when an animal has learned about warning coloration, negative experiences or punishment stimuli (such as inedibility) can have a stronger and more lasting effect than positive experiences. Accordingly it should be expected that there would be an excess of mimics over models in proportion to the degree of predominance of negative experiences over positive onces. When the deceptive animal possesses an innate reaction to the model the number of mimics can be virtually unlimited. 
	\item \textsl{``that the imitators differ from the bulk of their allies".} The most remarkable cases of mimicry are those where close relatives form divergent models. It is nevertheless highly possible that groups of closely related species may be similar mimics of the same model. Of course, it must always be remembered that species can look very different from their relatives because of adaptations which have nothing to do with mimicry. 
	\item \textsl{``that the imitation, however minute, is external and visible only, never extending to internal characters or to such as do not affect the external appearance."} This criteria, like the previous one is aimed at defining mimicry as a special adaptation. But if one of two closely related species both of which are poisonous and possess similar warning coloration, should secondarily lose its poisonous qualities, it would become a mimic and yet resemble its model in almost all its non-mimetic characters.
\end{itemize}

@@

\section{Mullerian Mimicry}
Bates was not able to explain some phenomena of mimicry. Occasionally two inedible unrelated butterfly species are amazingly similar in appearance. An explanation for this was provided by Fritz Muller in 1878. Like Bates, Muller observed and caught butterflies in Brazil. When there are multiple inedible species it is hard for predators to recognize each of them to know which one to consume and which one to avoid. Because of predator's limited memory, all these species still lose their number even after being inedible. So to save this loss, and to prevent more sacrifice of their own kind, inedible species from different family also tend to evolve to have similar appearance. This phenomena is referred to as Mullerian Mimicry in the name of Fritz Muller.

@@

\textsl{``According to punctuated equilibrium, therefore, evolution is not more of less both gradual and continual, as has been supposed for most purposes in most post-Darwinian theories. Rather, most morphological change is held to take place within geologically short periods, separated by vastly longer periods of comparative stasis \cite{eldredge1972}. The rapid, large changes all occur during speciation, that is to say during the branching of the evolutionary tree (cladogenesis);" \cite{turner1988}}

@@

\begin{itemize}
	\item If we imagine two different rings of similar warning patterns in one single habitat, similar enough that the envelop of protection afforded to one gives some protection to the more extreme variant of the other (in a simplified language, suppose that predators that have sampled species A sometimes avoid the more extreme members of species B because they mistake them as A), then these two species are subject to natural protection for mutual convergence of their patterns. Eventually they will form a rather accurate Mullerian mimicry ring.
	\item By contrast, if the two species have patterns so dissimilar that predators encountering one never imagine that it might be the other (the envelop of protection do not overlap the other species' phenotype distribution), then there is no selection whatever for convergence, and the two patterns will remain distinct indefinitely.
 \item But mutations of color patterns are occurring all the time. Suppose that species B happens to produce a mutation whole pattern falls within the envelop of protection afforded to species A. The mimicry need not be perfect, but if it is good enough, the mutation will have an advantage and will spread. In this way species B may, by a single mutation, become a Mullerian mimic of species A; in other word species B may switch from one Mullerian mimicry ring to another. Clearly, if the mimicry is not perfect, the two patterns will then gradually converge, making the resemblance ever closer. 
\end{itemize}

>>>>

%Chapter describing the mimetic process
\chapter{The Inspiration: Mimicry}
\label{chapter:mimicry}
\begin{quote}
\textsl{``It is not the strongest of the species that survives, nor the most intelligent that survives. It is the one that is the most adaptable to change." - Charles Darwin}
\end{quote}

@@

\begin{itemize}
	\item \textsl{``that the imitative species occur in the same area and occupy the same station as the imitated".} This increases the probability that a predator will affect both parties concerned. Nevertheless, the model could live in Africa and the mimic in Europe, or vice versa, if the deceptive predator were a migratory bird.
	\item \textsl{``that the imitators are always the more defenseless".} On the contrary, for the case of Mertensian mimicry among coral snakes the more offensive party may occasionally be the mimic.
	\item \textsl{``that the imitators are always less numerous in individuals".} What is meant here is that the deceived animal should encounter the mimic less often than the model. This is true when the deceived animal must learn to recognize model and mimic and when positive and negative experiences carry equal weight. But it is known from various learning experiments that when an animal has learned about warning coloration, negative experiences or punishment stimuli (such as inedibility) can have a stronger and more lasting effect than positive experiences. Accordingly it should be expected that there would be an excess of mimics over models in proportion to the degree of predominance of negative experiences over positive onces. When the deceptive animal possesses an innate reaction to the model, the number of mimics can be virtually unlimited. 
	\item \textsl{``that the imitators differ from the bulk of their allies".} The most remarkable cases of mimicry are those where close relatives form divergent models. It is nevertheless highly possible that groups of closely related species may be similar mimics of the same model. Of course, it must always be remembered that species can look very different from their relatives because of adaptations which have nothing to do with mimicry. 
	\item \textsl{``that the imitation, however minute, is external and visible only, never extending to internal characters or to such as do not affect the external appearance."} This criteria, like the previous one is aimed at defining mimicry as a special adaptation. But if one of two closely related species both of which are poisonous and possess similar warning coloration, should secondarily lose its poisonous qualities, it would become a mimic and yet resemble its model in almost all its non-mimetic characters.
\end{itemize}

@@

\section{Mullerian Mimicry}
\label{sec:mullerian-mimicry}
Bates was not able to explain some phenomena of mimicry. Occasionally two inedible unrelated butterfly species are amazingly similar in appearance. An explanation for this was provided by Fritz Muller in 1878. Like Bates, Muller observed and caught butterflies in Brazil. When there are multiple inedible species it is hard for predators to recognize each of them to know which one to consume and which one to avoid. Because of predator's limited memory, all these species still lose their number even after being inedible. So to save this loss, and to prevent more sacrifice of their own kind, inedible species from different family also tend to evolve to have similar appearance. This phenomena is referred to as Mullerian Mimicry in the name of Fritz Muller.

@@

\textsl{``According to punctuated equilibrium, therefore, evolution is not more or less both gradual and continual, as has been supposed for most purposes in most post-Darwinian theories. Rather, most morphological change is held to take place within geologically short periods, separated by vastly longer periods of comparative stasis \cite{eldredge1972}. The rapid, large changes all occur during speciation, that is to say during the branching of the evolutionary tree (cladogenesis);" \cite{turner1988}}

@@

\begin{itemize}
	\item If we imagine two different rings of similar warning patterns in one single habitat, similar enough that the envelop of protection afforded to one gives some protection to the more extreme variant of the other (in a simplified language, suppose that predators that have sampled species A sometimes avoid the more extreme members of species B because they mistake them as A), then these two species are subject to natural protection for mutual convergence of their patterns. Eventually they will form a rather accurate Mullerian mimicry ring.
	\item By contrast, if the two species have patterns so dissimilar that predators encountering one never imagine that it might be the other (the envelop of protection do not overlap the other species' phenotype distribution), then there is no selection whatever for convergence, and the two patterns will remain distinct indefinitely.
 \item But mutations of color patterns are occurring all the time. Suppose that species B happens to produce a mutation whose pattern falls within the envelop of protection afforded to species A. The mimicry need not be perfect, but if it is good enough, the mutation will have an advantage and will spread. In this way species B may, by a single mutation, become a Mullerian mimic of species A; in other word species B may switch from one Mullerian mimicry ring to another. Clearly, if the mimicry is not perfect, the two patterns will then gradually converge, making the resemblance ever closer. 
\end{itemize}

### 2010-12-13 docs/tex/result.tex
Chapter labels added...
Revision updates...

% All the result and analysis
\chapter{The Results}

\begin{quote}
\textsl{``The most incomprehensible thing about the world is that it is comprehensible." - Albert Einstein}
\end{quote}

@@

The plot in Figure \ref{fig:plot-2-prey} is simulation time verses prey population after running it for 10000 iterations. With the initial configuration in the above table we can observe that multiple rings of prey population has been created. Two prey species are considered to be in a ring if their CA pattern have a hamming distance within 10 bits. This parameter is mentioned in Table \ref{tab:ring-report-control-parameters}. Population of palatable species has been represented with line curve while population of unpalatable species has been presented with dotted curve. Different signs of squares, triangles and diamonds have been used to distinguish between species of prey population. The simulation was initiated with two prey species having CA rule of 30 and 110 and being palatable and unpalatable consecutively.

@@

We can observe from the above result that the evolution of mimicry has taken effect. A population of mimics were successfully able to exceed the population of other prey species, and the reason being avoidance by predators of prey pattern similar to unpalatable ones. We can conclude that Batesian mimicry has taken effect in the simulation.

@@

With the above set of parameters we receive the plot of Figure \ref{fig:plot-6-prey}. An enormous diversity of species can be observed from this experiment. In addition to the six prey species with which the simulation starts, the total number of mimicry rings reach nearly 50. Observing the top 8 as presented in the plot of figure \ref{fig:plot-6-prey}, effects of Batesian mimicry is obvious. Unpalatable CA Rule 190 and 30 dominates the population. Similarly unpalatable CA Rule 105 is also dominant. But after 10000 iterations CA Rule 55 seems to have higher population than 105 even though 55 is a \textsl{palatable} set of species. This anomaly could be the reason for storing large number of patterns in the predator Hopfiled memory. As we know that the number of patterns stored in Hopfiled network is inversely proportional to its recognition rate. In this simulation memory size is controlled by the 'Maximum Memory Configuration' which is set to 10 in table \ref{tab:config-table-6-prey}. Also it seems from this experiment that the 'Minimum Memory Configuration' being 6 is rather a large number. So there could be possibility of wrong recognition by Hopfield memory. Then again this phenomenon is common if we consider behavior of predators in nature. It is definitely hard for a single predator to memorize multiple poisonous patterns and because of which Mullerian mimicry takes an effect as mentioned in section \ref{sec:mullerian-mimicry}.

@@

The results according to figure \ref{fig:plot-4-prey-unp} are much expected. The population of unpalatable species have prevailed. After nearly 8000 iterations we can see unpalatable species of CA rule 55, 110, 30 and 190 have prevailed with the most population. All of their palatable counter parts are also increasing their population deceiving the predators. We observe the effect of the simulation up to 8000 iterations instead of 10000 as the total number of prey species increases at an enormous rate in this simulation within very short period of time. As most of the species are unpalatable, predators most do not get the opportunity to consume, causing a much lower prey death rate.

@@

This experiment is an ideal scenario for observing Mullerian mimicry. As mentioned in section \ref{sec:mullerian-mimicry}, Mullerian mimicry occurs between multiple species of unpalatable prey population. From the second work of Frank and Noble \cite{franks2003} mentioned in section \ref{subsec:models-by-frank-and-noble}, we note that multiple Mullerian mimicry rings are expected to converge into one large ring through the evolutionary process of punctuated equilibrium. But is this experiment as the predator's `Minimum Memory Configuration' is set to four, all predators have the capability to recognize four prey patterns before starting to make intelligent decision of consuming them. By setting this property to one and also increasing `Predator Demise Age' to 7000 while decreasing `Predator Age Interval' to 1500, we get the following plot.

>>>>

% All the result and analysis
\chapter{The Results}
\label{chapter:results}
\begin{quote}
\textsl{``The most incomprehensible thing about the world is that it is comprehensible." - Albert Einstein}
\end{quote}

@@

The plot in Figure \ref{fig:plot-2-prey} is simulation time verses prey population after running it for 10000 iterations. With the initial configuration in the above table we can observe that multiple rings of prey population have been created. Two prey species are considered to be in a ring if their CA pattern have a hamming distance within 10 bits. This parameter is mentioned in Table \ref{tab:ring-report-control-parameters}. Population of palatable species has been represented with line curve while population of unpalatable species has been presented with dotted curve. Different signs of squares, triangles and diamonds have been used to distinguish between species of prey population. The simulation was initiated with two prey species having CA rule of 30 and 110 and being palatable and unpalatable consecutively.

@@

We can observe from the above result that the evolution of mimicry has taken effect. A population of mimics were successfully able to exceed the population of other prey species, and the reason being, avoidance by predators of prey pattern similar to unpalatable ones. We can conclude that Batesian mimicry has taken effect in the simulation.

@@

With the above set of parameters we receive the plot of Figure \ref{fig:plot-6-prey}. An enormous diversity of species can be observed from this experiment. In addition to the six prey species with which the simulation starts, the total number of mimicry rings reach nearly 50. Observing the top 8 as presented in the plot of figure \ref{fig:plot-6-prey}, effects of Batesian mimicry is obvious. Unpalatable CA Rule 190 and 30 dominates the population. Similarly unpalatable CA Rule 105 is also dominant. But after 10000 iterations CA Rule 55 seems to have higher population than 105 even though 55 is a \textsl{palatable} set of species. This anomaly could be the reason for storing large number of patterns in the predator Hopfiled memory. As we know that the number of patterns stored in Hopfiled network is inversely proportional to its recognition rate. In this simulation memory size is controlled by the `Maximum Memory Configuration' which is set to 10 in table \ref{tab:config-table-6-prey}. Also it seems from this experiment that the `Minimum Memory Configuration' being 6 is rather a large number. So there could be possibility of wrong recognition by Hopfield memory. Then again this phenomenon is common if we consider behavior of predators in nature. It is definitely hard for a single predator to memorize multiple poisonous patterns and because of which Mullerian mimicry takes an effect as mentioned in section \ref{sec:mullerian-mimicry}.

@@

The results according to figure \ref{fig:plot-4-prey-unp} are much expected. The population of unpalatable species have prevailed. After nearly 8000 iterations we can see unpalatable species of CA rule 55, 110, 30 and 190 have prevailed with the most population. All of their palatable counter parts are also increasing their population deceiving the predators. We observe the effect of the simulation up to 8000 iterations instead of 10000 as the total number of prey species increases at an enormous rate in this simulation within very short period of time. As most of the species are unpalatable, predators mostly do not get the opportunity to consume, causing a much lower prey death rate.

@@

This experiment is an ideal scenario for observing Mullerian mimicry. As mentioned in section \ref{sec:mullerian-mimicry}, Mullerian mimicry occurs between multiple species of unpalatable prey population. From the second work of Frank and Noble \cite{franks2003} mentioned in section \ref{subsec:models-by-frank-and-noble}, we note that multiple Mullerian mimicry rings are expected to converge into one large ring through the evolutionary process of punctuated equilibrium. But in this experiment as the predator's `Minimum Memory Configuration' is set to four, all predators have the capability to recognize four prey patterns before starting to make intelligent decision of consuming them. By setting this property to one and also increasing `Predator Demise Age' to 7000 and decreasing `Predator Age Interval' to 1500, we get the following plot.
